### 2023-05-22 paper/my_flowchat.tex
modify contains and figures

\documentclass{article}
\usepackage[showframe]{geometry}
\usepackage{tikz}
\usepackage{bm}
\usetikzlibrary{
  arrows.meta,
  calc,
  fit,
  positioning,
  quotes,
  graphs,
  shapes.geometric
}
\begin{document}
% \makeatletter
\tikzset{
  meta box/.style={
    draw,
    black,
    very thick,
    text centered,
  },
  punkt/.style={
    meta box,
    rectangle,
    rounded corners,
    inner sep=.25em,
    minimum height=2em,
    minimum width=4em,
    align=center,
    text width=10em
  },
  round/.style={
    meta box,
    circle,
    minimum size=0,
    inner sep=0pt, 
    outer sep=0pt
  },
  every fit/.style={
    draw,
    thick,
    dashed,
    gray,
    inner xsep=.5em,
    inner ysep=.75em
  }
}
% helper macro
% \newcommand\tikz@expand@dimen[2]{\tikzset{minimum #2={#1/4}}}
% tikz styles: adjust min width/height of nodes
\tikzset{
  add dimen/.code 2 args={%
    \pgfkeysgetvalue{/pgf/minimum #1}\tikz@dimen@min
    \expandafter\tikz@expand@dimen\expandafter{\tikz@dimen@min + #2 * 2em}{#1}%
  },
  wider/.style={add dimen={width}{#1}},
  higher/.style={add dimen={height}{#1}},
}
% \makeatother

\begin{tikzpicture}[font=\small, node font=\small, node distance=1.5em]
    \node[punkt]     (input) {Input: RBS structure};
    \node[punkt, below=0.3of input] (get_SP) {
      get $SP$s by Eqs. 1 and 2 and Dijkstra Search
      };

    \node[punkt, below=0.3of get_SP] (get_A) {get $\bm{A}$ from Eq.1};
    \node[punkt, below=0.3of get_A] (get_Nset) {init $N_{set}=N_b$ };
    \node[punkt, below=0.3of get_Nset] (get_cb) {get $c_b$s by combinating $N_{set}$ batteries from $N_b$};
    \node[draw, diamond, aspect=2, below=0.3of get_cb] (is_check_all_cb) {are all $c_b$s checked?};
    \node[draw, diamond, aspect=2, right=1of is_check_all_cb] (is_Nset_converged) {is $N_{set}$ converged?};
    \node[punkt, above=of is_Nset_converged] (reset_Nset) {reset $N_{set}$ by dichotomy};
    \node[punkt, text width=15em, below=0.3of is_check_all_cb] (get_Xs) {
      select an unchecked $c_b$, and get its $\bm{X}_m$ by \\ 
      if switch $j$ $\in \bigcup_{i\in c_b}SP_i$:\\
             $\bm{X}[j]=1$ else $0$};
    \node[punkt, below=0.3of get_Xs] (get_Yn) {get $\bm{Y}_n$ by Eq.3};
    \node[draw, diamond, aspect=2, below=0.3of get_Yn] (is_Yn_invertible) {is $\bm{Y}_n$ invertible?};
    \node[punkt, right=1.3of is_Yn_invertible] (construct) {construct an effective solution};
    \node[punkt, below=0.3of is_Yn_invertible] (get_I) {get $I_o$ and $\bm{I}_b$ by Eq.4 and 5};

    \node[draw, diamond, aspect=2, below=0.3of get_I] (is_leq_Im) {is $\max \bm{I}_b \leq I_m$?};
    \node[punkt, right=1.3of is_leq_Im] (drop_eta) {drop this $\eta$};
    \node[punkt, below=0.3of is_leq_Im] (get_eta) {get $\eta$ by Eq.6};
    \node[punkt, below=0.3of get_eta] (update_max_eta) {update $\max \eta$};
    \node[punkt, right=1of is_Nset_converged] (output) {Output: $\max \eta$};

    \node[round,left=1.5of update_max_eta](point1){};

    \graph{
      (input) -> (get_SP) -> (get_A) -> (get_Nset) -> (get_cb) -> (is_check_all_cb) ->["No"] (get_Xs) -> (get_Yn) -> (is_Yn_invertible) ->["Yes"] (get_I) -> (is_leq_Im) ->["Yes"] (get_eta) -> (update_max_eta);
      (is_check_all_cb) ->["Yes"] (is_Nset_converged) ->["No"] (reset_Nset) -> (get_cb);
      (is_Yn_invertible) ->["No"] (construct) ->[to path={|- (\tikztotarget)}] (get_I);
      (is_leq_Im) ->["No"] (drop_eta) ->[to path={|- (\tikztotarget)}] (update_max_eta);
      (is_Nset_converged) ->["Yes"] (output);
      (update_max_eta) -- (point1) ->[to path={|- (\tikztotarget)}] (is_check_all_cb);
    };

\end{tikzpicture}
\end{document}