### 2023-09-12 paper/SST/SST_main.tex
完成了修改 (v1.2)

% 【done】 
% =====0909
% 按 Q 修改意见修改 0901 【优先】
% =====0910
% 1. 摘要修改，量化我们的成果：explain the most important achievements of the proposed method quantitatively
% 2. 介绍修改，相关文献：
%   2.1 RBS 结构分析相关的
%   2.2 网络流（最大流）相关（但因为问题不同，无法使用，提一句穷举）
% 3. 介绍修改，逐点列出我们的贡献：state research contribution and achievements as bullet points at the end of the Introduction section
% 4. 讨论修改，增加算法复杂度，以及和穷举法的比较（该领域尚为空白，没有sota的算法）
% 5. 讨论修改，增加算法的 limitations（根据复杂度，大规模不好用）
% 6. 讨论修改，增加结果的准确性介绍（和穷举结果比较）：explain about the accuracy, sufficiency and reliability of their results? How do they verify and validate the results
% 7. 修改算法的第15步
% 8. 重写总结
% 9. 添加文献
% =====0911
% 1. introduction 整体逻辑不流畅，跳跃过大
% 2. introduction 文献
% 2. conclusion 结尾，可以加上一两句说说这个方法的应用价值之类的，吹一吹
% 3. abstract
%   3.1 删去 75000 倍的效率；
%   3.2 凝练一点，就是说这个方法把遍历法中的枚举开关状态组合，转换为了对最短路径的组合；
%   3.3 算法复杂度、效率提升倍数=2^开关数/2^电池数
%   3.4 调整 discussion 的相关内容

% 【doing】

% 【todo】
% ==【14号前交给润色机构】

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\usepackage{setspace}
\usepackage[margin=1.25in]{geometry}
\usepackage{graphicx}
% \graphicspath{ {./figures/} }
\graphicspath{ {../figures/} }
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{lineno}
\linenumbers

\usepackage{enumerate}
\DeclareMathOperator{\diag}{diag}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{footnote}
\usepackage{bm}
\usepackage[ruled,linesnumbered]{algorithm2e}
\setcounter{MaxMatrixCols}{20}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, calc, fit, positioning, quotes, graphs, shapes.geometric}
\def\T{\mathrm{T}}

% \usepackage[defaultcolor=red]{changes}
% \added{xxx}
% \deleted{xxxx}
% \replaced{a}{b} 原文的b替换成a
\usepackage[final]{changes}
% \usepackage{easyReview} % texstudio 部分

%%%%%% Bibliography %%%%%%
% Replace "sample" in the \addbibresource line below with the name of your .bib file.
% \usepackage[style=nejm, citestyle=numeric-comp, sorting=none]{biblatex}
% \addbibresource{SST_main.bib}

%%%%%% Title %%%%%%
% Full titles can be a maximum of 100 characters, including spaces. 
% Title Format: Use title case, capitalizing the first letter of each word, except for certain small words, such as articles and short prepositions
\title{Maximum Allowable Current Determination of RBS By Using a Directed Graph Model and Greedy Algorithm}

%%%%%% Authors %%%%%%
% Authors should be listed in order of contribution to the paper, by first name, then middle initial (if any), followed by last name.
% Authors should be listed in the order in which they will appear in the published version if the manuscript is accepted. 
% Use an asterisk (*) to identify the corresponding author, and be sure to include that person’s e-mail address. Use symbols (in this order: †, ‡, §, ||, ¶, #, ††, ‡‡, etc.) for author notes, such as present addresses, “These authors contributed equally to this work” notations, and similar information.
% You can include group authors, but please include a list of the actual authors (the group members) in the Supplementary Materials.
\author[1$\dag$]{Binghui Xu}
\author[1$\dag$]{Guangbin Hua}
\author[1*]{Cheng Qian}
\author[1,2]{Quan Xia}
\author[1]{Bo Sun}
\author[1]{Yi Ren}
\author[1]{Zili Wang}

%%%%%% Affiliations %%%%%%
\affil[1]{School of Reliability and Systems Engineering, Beihang University, Beijing, 100191, China}
\affil[2]{School of Aeronautic Science and Engineering at Beihang University, Beijing, China}
\affil[*]{Address correspondence to: cqian@buaa.edu.cn}
\affil[$\dag$]{These authors contributed equally to this work.}

%%%%%% Date %%%%%%
% Date is optional
\date{}

%%%%%% Spacing %%%%%%
% Use paragraph spacing of 1.5 or 2 (for double spacing, use command \doublespacing)
\onehalfspacing

\begin{document}

\maketitle

%%%%%% Abstract %%%%%%
\begin{abstract}
Reconfigurable Battery Systems (RBSs) present a promising alternative to traditional battery systems due to their flexible and dynamically changeable topological structure subjected to battery charging and discharging strategies.
During the operation of the RBS, the Maximum Allowable Current (MAC) of system that ensures each battery's current remains within a safe range, is a critical indicator to guide the system's reconfiguring control, ensuring its safety and reliability. 
In this paper, we firstly propose a calculation method for the MAC of arbitrary RBS using a greedy algorithm in conjunction with a directed graph model of the RBS.
\added{
By introducing the shortest path of the battery, the greedy algorithm transforms the enumeration of switch states in the brute force algorithm into the combination of the shortest paths, which greatly increases the efficiency of the determination of the MAC.
The directed graph model, based on the equivalent circuit, provides a specific method for calculating the MAC of a given structure.
This method has been validated on two published 4-battery-RBSs and one with more complex structure.
It obtaining the same results as the brute force algorithm method, while achieves significantly improved computational efficiency ($N_s 2^{N_s - N_b} \log N_b$ times faster than the brute force algorithm for a RBS with $N_b$ batteries and $N_s$ switches, theoretically).
The main advantage of this method is its ability to calculate the MAC of RBSs with arbitrary structures, even in the scenarios with random isolated batteries.
}
\deleted{
In this method, a new directed graph model is developed to model the structure of RBS, and a greedy algorithm is designed to find the possible circuit that enable MAC.
Then, the MAC is calculated based on the circuit in-cooperate with the equivalent model of batteries and switches. 
The effectiveness of the proposed method is validated by a novel and complex RBS structure. 
The results show that this method is capable to calculated the MAC of RBSs with different structures or different battery sizes efficiently, which proves the correctness of this method and its potential in facilitating next-generation RBS designs and applications, including battery isolation.
}
\end{abstract}

%%%%%% Main Text %%%%%%

\section{Introduction}

Battery Energy Storage Systems (BESSs) are extensively employed in various applications, such as wind power plants and space power systems, to store and release high-quality electrical energy \cite{desiqueiraControlStrategySmooth2021,yangBatteryEnergyStorage2018,choCommercialResearchBattery2015,zhangDevelopmentProspectChinese2021,schwanbeckInternationalSpaceStation2019}.
Typically, a BESS consists of numerous batteries interconnected by series-parallel circuitry to provide the required capacity storage.
However, traditional BESSs, in which the batteries are connected in a fixed topology, exhibit a significant weakness in their worst battery due to the so-called cask effect.
Moreover, if this worst battery fails \replaced{from}{during} operation, it can exacerbate the degradation of other batteries with a high possibility, leading to reliability and safety issues \cite{yangUnbalancedDischargingAging2016,fengPropagationMechanismsDiagnosis2019,jeevarajanBatterySafetyQualifications2012}.
These problems have become significant technical barriers in the development of new-generation space vehicles \replaced{unfortunately}{and urgently need to be addressed} \cite{pomboHybridPowerSystem2021}.

Reconfigurable Battery System (RBS), which can dynamically switch to different circuit topology configurations as required, is expected to solve the above problems\cite{hanNextGenerationBatteryManagement2020a}.
The ability of switching circuit helps to isolate unhealthy batteries, \replaced{thereby improving}{and thereby improve} the safety and reliability of the battery system.
\replaced{In order to illustrate the working principle of RBS, a typical RBS structure developed by Visairo \cite{visairoReconfigurableBatteryPack2008}(Figure \ref{fig:stru-Visairo}) is taken as an example to show the reconfiguration process.}{Figure \ref{fig:stru-Visairo} shows a typical RBS structure developed by Visairo \cite{visairoReconfigurableBatteryPack2008} for dynamically adjusting the output voltage and current.}
In this structure, the batteries can be connected not only in series when the switches $S_1$, $S_5$, $S_6$, $S_7$, $S_8$, $S_9$, and $S_{13}$ are closed (Figure \ref{fig:stru-Visairo-serial}), but also in parallel when $S_1$, $S_2$, $S_3$, $S_4$, $S_5$, $S_9$, $S_{10}$, $S_{11}$, $S_{12}$, and $S_{13}$ are closed (Figure \ref{fig:stru-Visairo-parallel}).
Furthermore, when an unhealthy battery, for instance the orange one $B_3$ in Figure \ref{fig:stru-Visairo-isolate}, appears in the RBS, it can be isolated by opening its two adjacent switches (i.e. $S_4$ and $S_{11}$), ensuring the system still remains a reliable working mode.

\begin{figure}[htbp]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{stru-V-origin.png}
        \caption{}
        \label{fig:stru-Visairo}
    \end{subfigure}
    \hspace{0.05\textwidth}
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{stru-V-serial.png}
        \caption{}
        \label{fig:stru-Visairo-serial}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{stru-V-parallel.png}
        \caption{}
        \label{fig:stru-Visairo-parallel}
    \end{subfigure}
    \hspace{0.05\textwidth}
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{stru-V-isolate.png}
        \caption{}
        \label{fig:stru-Visairo-isolate}
    \end{subfigure}
    \caption{
        (a) The RBS structure proposed by Visairo\cite{visairoReconfigurableBatteryPack2008}, with
        all batteries in (b) series connection, (c) parallel connection, and
        (d) battery $B_3$ isolated.
        }
    \label{fig:arch}
\end{figure}

\added{
% 目前，许多研究人员利用 RBS的灵活地可重构特点，设计多种结构，以实现并改善系统的 performance
% Ci【139】等人提出的结构能够动态调整电池放电倍率以充分挖掘各电池的可用容量；
% Jan【149，150】等人针对电动汽车充电时不断变化的电压需求，提出了一种能够动态调整电池串联数的结构。
% Visairo 【140】等人的结构可以根据负载条件改变系统的输出电压，从而减少供电过程中电压调节器的功率损耗以提升能量利用率；
% 同样处于提升系统能量效率的目的，Lawson等人\cite{lawsonSoftwareConfigurableBattery2012}和He等人【148】也提出过更为简化的结构。Kim等人【143】则通过向结构引入多端口，提升了系统对电池故障的恢复能力。
Currently, many researchers utilize the flexibility and reconfigurability of RBS to design various structures in order to meet application requirements and improve system performances. 
Ci et al. \cite{ci2007novel} proposed a RBS structure that can dynamically adjust the batteries discharge rate to fully exploit the available capacity of each battery. 
Jan's \cite{9209774,engelhardt2021double} structures had the ability to reconfigure structures with variant batteries in series to reach the voltage requirements during electric vehicle charging, which constantly change.
The structure proposed by Visairo et al.\cite{visairoReconfigurableBatteryPack2008}, as shown in Figure \ref{fig:stru-Visairo}, can change the system's output voltage based on load conditions, thereby reducing the power loss of the voltage regulator during the power supply process and improving energy utilization efficiency. 
For the same purpose to enhance the energy efficiency of the system, Lawson et al. \cite{lawsonSoftwareConfigurableBattery2012} and He et al. \cite{he2014reconfiguration} also proposed simplified structures which has less switches than Visairo's design.
Kim et al. \cite{kim2009dynamic} improved the system's ability to recover from battery failures by introducing multiple ports to the structure. 
% 这些结构的提出很大程度上依赖研究人员的经验，并且缺少严谨的结构分析方法和客观全面的评价指标。
% 这可能会导致结构设计不够优化，无法充分发挥RBS的性能优势；甚至存在潜在的安全隐患，无法让RBS投入实际应用。
% To tackle these issues, some analyses on the RBS structure and performance have been conducted. 
% For instance, Chen et al. \cite{chenSneakCircuitTheory2021} proposed a systematic approach based on the sneak circuit theory is proposed to fundamentally avoid the short-circuit problem of reconfigurable battery systems.
% This research goes through and analyzes all paths between the positive and negative terminals of each battery in the RBS, and identifies paths that only contain switches as short-circuit paths to form a short-circuit path table for pre-check and before system reconfiguration.
% Ham【162】等人则针对一种特定的RBS结构，推导出 maximum switch current during battery system reeconfiguration 的解析表达式。
% 该工作可用于指导RBS的开关选型，为RBS的硬件设计提供支撑。
The design of these structures mainly relies on the experience of researchers and lacks rigorous structural analysis methods and quantitative evaluation criteria. 
This may result in suboptimal structures, inability to fully leverage the performance advantages of RBS, and even potential safety hazards that prevent practical applications of RBS. 
To tackle these issues, some analyses on the RBS structure and performance have been conducted. 
For instance, Chen et al.\cite{chenSneakCircuitTheory2021} proposed a systematic approach based on the sneak circuit theory to fundamentally avoid the short-circuit problem of reconfigurable battery systems. 
This research thoroughly analyzes all paths between the positive and negative terminals of each battery in the RBS and identifies paths that only contain switches as short-circuit paths, forming a short-circuit path table for pre-checking before system reconfiguration. 
Han et al.\cite{han2021analysis} derived an analytical expression for the maximum switch current during battery system reconfiguration for a specific RBS structure. 
This helps guide the selection of switches and provide support for the hardware design of RBS.
}

\added{
% 当前，对于RBS许用电流的分析仍是空白。
% 电池具有合适的工作电流范围，过大的电流会导致电池的性能下降和衰退，甚至损坏电池。【文献】
% 通过有限数量的电池组成的RBS也同样存在系统输出电流的上限，即 maximum allowable current (MAC)。
% 超过该阈值会导系统内部部分电池的电流过载。
% 因此，需要一种能够确定系统 MAC 的 RBS结构的电流分析方法和求解算法
There is still a blank of analysis the allowable current of RBS.
The battery has an appropriate range of operating currents, and excessive current can lead to a decrease and degradation in battery performance, or even damage\cite{lotfi2017li}.
For RBS composed of finite batteries, there is also an upper limit on the system's output current, known as the maximum allowable current (MAC). 
Exceeding this threshold can result in current overload on certain batteries of the system. 
Therefore, there is a requirement for analysis method on RBS current and solving algorithm that can determine the system's MAC.
The maximum flow problem (MFP) in graph theory has a similar formulation to the RBS MAC determination.
MFP is defined as finding the path with the maximum flow in a network flow with a single source node and a single sink node.
In the RBS, the system positive and negative terminals can be taken as the source and sink nodes respectively, and the current can be regarded as the flow of the graph. 
Meanwhile, the Kirchhoff's laws that describe the current in circuit, and the function of switches to control the circuit on-off should be appropriately translated into constraints in the MFP.
This strongly increases the complexity and makes it a special maximum flow problem that never been studied before.
Despite mature algorithms for the MFP, such as the Edmonds-Karp algorithm \cite{10.1145/321694.321699} and Dinic's blocking flow algorithm \cite{doi:10.1137/0204043}, these algorithms cannot be directly applied to solve the MAC due to the speciality of the problem. 
Moreover, these MFP algorithms need to comprehensively compute all possible reconfigured structures to obtain the MAC, which means highly inefficience and not suitable for large-scale RBS.
}

\deleted{
The complex connection structure between batteries and switches in the RBS provides flexibility but also introduces challenges in design and operational control. 
Unlike traditional BESSs with fixed outputs, the RBS output must be dynamically adjusted by controlling switch states to meet external load requirements. 
This necessitates additional, time-consuming output performance analysis during design and corresponding control strategies. 
An incorrect switch control strategy may cause battery short-circuiting or overload, risking the entire system. 
The Maximum Allowable Current (MAC), an RBS performance indicator, can guide designers in addressing this issue. 
MAC is defined as the maximum RBS output current that ensures each battery's current remains within a safe range.
Therefore, it provides a benchmark for RBS output current, protecting individual batteries and identifying overall system output limits during operation. 
Despite its importance, no method currently exists for automatically evaluating MAC for RBSs.
In particular, when one or more random cells are isolated, there is still no method to determine the MAC of the remaining RBS in time to assist the system in adjusting the control strategy timely. 
A universal and automatical method for calculating RBS MAC is urgently needed for practical applications.
}

\added{
To effectively solve the switches states of MAC while strongly decreasing the unnecessary reconfigured structures to compute, and still robust in the scenarios with battery isolation, the greedy algorithm is adopted and adapted to solve the MAC problem in this paper.
Meanwhile, an improved direct graph model, considering the voltage, internal resistance, and maximum allowable current of the battery, and the external load, is porposed and emploied to obtain the accuracy value of MAC.
The correctness and effectiveness of the method in this paper are verified by two published 4-battery-RBSs and one with more complex structure.
}
\deleted{In this study, a directed graph model and greedy algorithm are employed to determine the MAC of RBS and the corresponding control strategy, effectively calculating the MAC for RBSs with arbitrary structures, including scenarios with isolated batteries.}
\added{The main contributions of the paper can be summarized as follows:}
\begin{itemize}
  \item \added{Application of the greedy algorithm to solving the MAC of a given RBS for the first time.}
  \item \added{Presenting a improved directed graph model to analyse the current of RBS, which considers the voltage, internal resistance, and maximum allowable current of the battery, and the external load.}
  \item \added{Proposing a new RBS structure and comparing it with two proposed RBS structures in the scenario with randomly isolated batteries.}
\end{itemize}

The remainder of this paper is organized as follows: 
Section II presents the framework and details of the proposed directed graph model and the greedy algorithm. 
Section III demonstrated a case study of using the proposed method to determine the MAC of a novel and complex structure. 
The calculation results \added{, the algorithm's computational complexity} and scenarios such as batteries isolation also are discussed. 
Finally, the concluding remarks are drawn in Section IV.

\section{Methodology}

The central principle of this method is to make the batteries in RBS connected in parallel as much as possible, thereby maximizing the output current of the RBS.
To universally and automatically achieve this, the overall process is divided into four steps, as shown in Figure \ref{fig:main}.
Firstly, a directed graph model is established for subsequent \replaced{computations.}{computing,} \replaced{The model}{, which} not only contains the connected relationships between batteries and switches, but also retains the performance parameters of the batteries.
Subsequently, based on the equivalent circuit, the MAC problem is transformed into specific objective functions and constraints.
Then, the shortest paths (\replaced{$SP$}{SP}s, where additional batteries and switches on the path are penalized as distance) for the batteries are obtained using the Dijkstra algorithm to guide the batteries in the RBS connect in parallel.
Finally, a greedy algorithm is employed to organize the switches, allowing the batteries to connect via their \replaced{$SP$}{SP}s while satisfying the constraints, resulting in the MAC of the RBS.

\begin{figure}[htbp]
    \centering
    \begin{subfigure}[b]{0.8\textwidth}
        \includegraphics[width=\textwidth]{main.png}
    \end{subfigure}
    \caption{ 
        Diagram of this method, which contains four main steps.
    }
    \label{fig:main}
\end{figure}

\subsection{Directed graph Model}

\begin{figure}[htbp]
    \centering
    \begin{subfigure}[b]{0.31\textwidth}
        \includegraphics[width=\textwidth]{direct-graph-he.png}
        \caption{}
        \label{fig:direct-graph-he}
    \end{subfigure}
    \hspace{0.02\textwidth}
    \begin{subfigure}[b]{0.23\textwidth}
        \includegraphics[width=\textwidth]{direct-graph-xu.png}
        \caption{}
        \label{fig:direct-graph-xu}
    \end{subfigure}
    \hspace{0.02\textwidth}
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{direct-graph-my.png}
        \caption{}
        \label{fig:direct-graph-my}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.8\textwidth}
        \includegraphics[width=\textwidth]{battery_simple.png}
        \caption{}
        \label{fig:battery_simple}
    \end{subfigure}
    \caption{ 
        Directed graph models used in (a) He's work \cite{heExploringAdaptiveReconfiguration2013}, (b) our previous work , and (c) \added{the improved model in} this paper.
        (d) The equivalent circuit of a battery in this method.
    }
\end{figure}

He et al. \cite{heExploringAdaptiveReconfiguration2013} once proposed an abstracted directed graph model for RBS, where the nodes represented the batteries, the edges represented the configuration flexibility, and the weight of each vertex corresponded to the battery voltage ( Figure \ref{fig:direct-graph-he}). 
The model effectively captured all potential system configurations and offered a direct metric for configuration flexibility, but it did not specify the physical implementation of the connectivity between batteries, meaning one graph might \replaced{correspond to}{have had} multiple RBS structures.
We previously proposed a novel directed graph model that, \replaced{completely different from}{in contrast to} He's model, used nodes to represent the connections between batteries and switches, and directed edges to represent batteries and switches (Figure \ref{fig:direct-graph-xu}), allowing for a one-to-one correspondence between the RBS structure and the directed graph model. 
This model was able to accurately and comprehensively represent the RBS topological structure but could not be used for quantitative MAC calculations due to the lack of consideration \replaced{on}{for} \replaced{the voltage, internal resistance and maximum allowable current of the battery}{battery and switch performance parameters}. 
To address this, an \replaced{improvement on our previous model is conducted by}{improved directed graph model is used here based on our original model,} adding electromotive force and resistance attributes on the edges based \replaced{on its}{to} equivalent circuits \deleted{(Figure \ref{fig:direct-graph-my})}.
The model also considers the external load as an equivalent resistance and integrate it into the analysis, making it a complete circuit model for later circuit analysis.
\added{Figure \ref{fig:direct-graph-my} shows the improved directed graph model used in this paper.}
The following will provide a detailed explanation of the method for equating the components in RBS and constructing the directed graph model.

In order to use circuit analysis methods to solve the MAC of the RBS, the components in the RBS are equated to ideal circuit elements.
\added{For instance,} \replaced{a}{A}s shown in Figure \ref{fig:battery_simple}, the battery in the RBS can be represented as a black-box circuit consisting of two resistors (i.e., $r_1$ and $r_2$) and a capacitor (i.e., $C$), known as the Thevenin model\cite{hongwenheStateofChargeEstimationLithiumIon2011,mousavig.VariousBatteryModels2014}.
With an emphasis on the stable output of the RBS, the capacitor in the Thevenin model can be considered as an open circuit without affecting the steady-state current.
Therefore, the battery $i$ in the RBS can be simplified as the series connection between a constant voltage source $u_{i}$ and a resistor $r_{i}$.
Furthermore, the state of switch $j$ in the RBS is represented by a binary variable $x_j$, where 0 is for ON and 1 is for OFF, respectively.
When the switch is closed, it can be regarded as a resistor with a very small resistance value $r_{j}$.
Lastly, the external load is considered as a resistor with a value of $R_o$.

For a given RBS structure, \replaced{its directed graph model $G(V,E)$ is constructed as follows:}{the directed graph model for the RBS is constructed as a directed graph $G(V,E)$ in such a way that:}
\begin{enumerate}
    \item Nodes:
        The nodes in the directed graph correspond to the connection points of components in the actual RBS. 
        Assuming there are a total of $N$ nodes in the RBS, for the sake of convenience, the anode of the RBS is denoted as $v_1$ and the cathode as $v_N$.
    \item Edges:
        The edges in the directed graph correspond to the batteries, switches, and external electrical loads in the actual RBS.
        Therefore, there are three types of directed edges. 
        For a battery $B_i$, its directed edge $e_i$ is drawn from the cathode to the anode, as the battery only allows current to flow in one direction when in operation.
        For a switch $S_j$, since it is allowed to work under bi-directional currents, it is represented by a pair of directed edges with two-way directions. 
        Regarding the external electronic load, as it is connected to the anode and cathode of the RBS, a directed edge from $v_N$ to $v_1$ is used to represent it. 
        In conclusion, for a given RBS structure with $N_b$ batteries and $N_s$ switches, the total number of directed edges is $N_b+2N_s+1$, where 1 refers to the external electrical load.
    \item Edges' attributes:
        Each edge is assigned two attributes, voltage difference and resistance, based on the equivalent method mentioned above.
        The values for the battery $B_i$, switch $S_j$, and external loads correspond to $(u_i, r_i)$, $(0, r_j)$, and $(0, R_o)$, respectively.
\end{enumerate}

\subsection{Constraints and Objective Function}

\replaced{For a given RBS, determination of its MAC}{Based on the definition of MAC, determining the MAC of RBS} involves maximizing the RBS output current while ensuring that the currents of all batteries do not exceed the batteries' maximum allowable current. 
In this subsection, the constraints and objective function to solve the RBS's MAC will be established through circuit analysis, based on \replaced{its directed graph model provided in the previous subsection}{the previously constructed directed graph model}.

First, the topology in the directed graph model is represented in matrix form $\bm{A}$, known as the incidence matrix , \replaced{defined}{to facilitate circuit analysis.
The specific definition of the incidence matrix is shown} in Equation \ref{eq:A}.
\begin{align}\label{eq:A}
    a_{kl}=
    \begin{cases}
        1,  & \text{edge $l$ leaves node $k$},\\
        -1, & \text{edge $l$ enters node $k$},\\
        0,  & \text{otherwise}.
    \end{cases}
\end{align}
For a directed graph consisting of $N$ nodes and $N_b+2N_s+1$ directed edges, its incidence matrix $\bm{A}$ is an $N\times(N_b+2N_s+1)$ matrix. 
In this matrix, the rows and columns represent the nodes and edges of the directed graph, respectively.
By distinguishing the components in the RBS corresponding to each column , $\bm{A}$ can be rewritten as:
\begin{equation}\label{eq:A_bso}
    \bm{A} =
    \begin{bmatrix}
        \bm{A}_b & \bm{A}_s & \bm{A}_o
    \end{bmatrix},
\end{equation}
where $\bm{A}_b$, $\bm{A}_s$ and $\bm{A}_o$ are the sub-matrices corresponding to the batteries, switches and external electrical load, respectively.
To alleviate computational complexity, matrix $\bm{A}$ undergoes dimensionality reduction.
Since each directed edge has one node to leave and one to enter, the sum of the values in every column of $\bm{A}$ is zero.
Therefore removing \replaced{the last}{any single one} row will not result in a loss of information. 
\deleted{Without loss of generality, the last row is removed here.}
On the other hand, since each switch in the RBS is represented by a pair of directed edges with two-way directions, the two columns corresponding to the switch are mutually opposite.
Thus, for the sub-matrix $\bm{A}_s$, only one column is retained for each pair of columns representing the same switch.
As a result, $\bm{A}$ can be reduced to a $(N-1)\times(N_b+N_s+1)$ matrix, denoted as $\bm{\tilde{A}}$, for further calculation of current and voltage.
Similar to Equation \ref{eq:A_bso}, $\bm{\tilde{A}}$ can be rewritten as:
\begin{equation}\label{eq:A_bso_tilde}
    \bm{\tilde{A}} =
    \begin{bmatrix}
        \bm{\tilde{A}}_b & \bm{\tilde{A}}_s & \bm{\tilde{A}}_o
    \end{bmatrix}.
\end{equation}

After obtaining the incidence matrix, the currents of all batteries and output in RBS are determined by solving the circuit equations.
According to Kirchhoffs law, we have
\begin{align}\label{eq:Kirchhoffs_law}
    \begin{cases}
        \bm{\tilde{A}} \bm{I} = \bm{0}, \\
        \bm{U}        = \bm{\tilde{A}}^\T \bm{U}_n,
    \end{cases}
\end{align}
where $\bm{I}$ and $\bm{U}$ indicate the current and voltage difference arrays of the $N_b+N_s+1$ edges, respectively;
$\bm{U}_n$ is the voltage array of the $N-1$ nodes.
These directed edges are treated as generalized branches and expressed in matrix form as follows
\begin{equation}\label{eq:generalized_branches}
    \bm{I} = \bm{Y}\bm{X} \bm{U} - \bm{Y}\bm{X} \bm{U}_s +\bm{I}_s,
\end{equation}
where $\bm{U}_s$ and $\bm{I}_s$ denote the source voltage and source current of the generalized branches, respectively.
Because all batteries have been equivalent to voltage sources rather than current sources in the previous subsection, all elements of the array $\bm{I}_s$ are 0, 
while the elements of the array $\bm{U}_s$ are equal to the first attribute of the corresponding edges in the directed graph.
The $\bm{Y}$ in \added{Equation} \ref{eq:generalized_branches} is the admittance matrix of the circuit, defined as the inverse of the impedance matrix.
\replaced{The}{That is the} elements \replaced{on}{of} the diagonal \added{of} matrix $\bm{Y}$ are equal to the reciprocal of \added{the resistance, which is} the second attribute of the corresponding edges in the directed graph, and the off-diagonal elements \added{$\bm{Y}$} are 0.
The $\bm{X}$ is the state matrix, which describes whether the RBS batteries and switches are allowed to pass current.
It is defined as
\begin{equation}\label{eq:X}
    \bm{X} = \diag(
    \underbrace{1, 0 \cdots, 1}_{N_b~\text{of}~0/1},
    \underbrace{1, 0 \cdots, 1}_{N_s~\text{of}~0/1},
    1)
    =\begin{bmatrix}
        \bm{X}_b & & \\
        & \bm{X}_s &\\
        & & 1
    \end{bmatrix}.
\end{equation}
Where the elements $x_i$ of the matrix $\bm{X}_b$ represent whether the battery $i$ has been removed from the circuit, with $x_i=1$ indicating removal and $x_i=0$ indicating that it is still available to supply power. 
When all batteries are health and capable of providing current to the external load, $\bm{X}_b$ is an identity matrix. 
The elements $x_j$ of the matrix $\bm{X}_s$ represent whether the switch $j$ is closed, with $x_j=1$ indicating closure and $x_j=0$ indicating disconnection, which is consistent with the previous subsection.

Theoretically, the output current $I_o$ and the currents of each battery $\bm{I}_b$ in the RBS  can be determined by solving Equations \ref{eq:Kirchhoffs_law}, \ref{eq:generalized_branches}, and \ref{eq:X} under any given state $\bm{X}$.
In order to obtain specific constraint conditions and objective functions, it is further assumed that all batteries have the same electromotive force and internal resistance, denoted as $u_b$ and $r_b$, respectively.
This allows for the derivation of explicit expressions for $I_o$ and $\bm{I}_b$.
After derivation and simplification, the output current $I_o$ and the currents of each battery $\bm{I}_b$ are ultimately represented as Equations \ref{eq:I_o} and \ref{eq:I_b}, respectively.
\begin{equation}\label{eq:I_o}
    I_o = \frac{1}{R_o r_b} \bm{\tilde{A}}_o^\T \bm{Y}_n^{-1}(\bm{X}) \bm{\tilde{A}}_b \bm{U}_b,
\end{equation}
\begin{equation}\label{eq:I_b}
    \bm{I}_b = \frac{1}{r_b^2}[\bm{\tilde{A}}_b^\T \bm{Y}_n^{-1}(\bm{X}) \bm{\tilde{A}}_b\bm{U}_b -r_b \bm{U}_b],
\end{equation}
where $\bm{U}_b$ is a $N_b\times 1$ array with all elements equaling to $u_b$;
$\bm{Y}_n$ is the equivalent admittance matrix of the circuit, defined as
\begin{equation}\label{eq:Yn}
    \bm{Y}_n (\bm{X}) = \frac{1}{R_o} \bm{\tilde{A}}_o\bm{\tilde{A}}_o^\T + \frac{1}{r_b} \bm{\tilde{A}}_b\bm{X}_b\bm{\tilde{A}}_b^\T + \frac{1}{r_s}\bm{\tilde{A}}_s\bm{X}_s\bm{\tilde{A}}_s^\T.
\end{equation}

To characterize the current output capacity of the RBS structure under different switching states, an indicator $\eta$ is defined by the ratio of $I_o$ and $\max (\bm{I}_b)$ shown in Equation \ref{eq:eta}:
\begin{equation}\label{eq:eta}
    \eta = \frac{I_o}{\max (\bm{I}_b)}.
\end{equation}
Finally the problem of solving MAC can be formulated as
\begin{align}
    & \max \eta(\bm{X}_s) \label{eq:max_eta}\\
    \text{s.t.} & \max (\bm{I}_b) \leq I_m, \label{eq:Ib_leq_Im}
\end{align}
where $I_m$ is the maximum allowable current of the battery.

However, it is \added{still} computationally difficult to solve \ref{eq:max_eta} because of the $\bm{Y}_n^{-1}$.
On one hand, due to the introduction of nonlinear terms by $\bm{Y}_n^{-1}$, many effective methods in linear optimization are not suitable for this \replaced{scenario}{problem}.
On the other hand, the rank of $Y_{n}$ is proportional to the number of batteries and switches, which can be very large for a large RBS system, leading to significant computational burden.
Therefore, intelligent algorithms that rely on evolving by iteration may face efficiency \replaced{problem}{issues} when dealing with large RBS system.
In order to address this issue, the problem should be considered from the perspective of guiding the RBS to reconstruct as many parallel structures as possible.
Consequently, a greedy algorithm based on the shortest path is proposed. 
The detailed implementation process is presented in the following two subsections.

\subsection{Shortest Path}

The path $p$ used in this method is defined as the complete route that passes through one battery (or a consecutive series of batteries) and closed switches, connecting the anode $v_1$ to the cathode $v_N$ of the RBS.
By applying a penalty to the series-connected batteries on the path, where additional batteries imply a longer distance, the algorithm encourages the RBS to form parallel structures as much as possible.
Meanwhile, to reduce the number of switches controlled during the reconstruction process, a penalty is also applied to the total number of switches on the path, while ensuring the minimum number of batteries.
Therefore, the distance $\omega$ of the path $p$ is defined by the following equation: 
\begin{equation}\label{eq:weight}
    \omega(p) = N_s \cdot n_b (p) + n_s (p),
\end{equation}
where $N_s$ is the total number of switches in the system; 
$n_b(p)$ and $n_s(p)$ are number of batteries and switches in the path $p$ respectively. 
Moreover, the shortest path $SP_i$ is defined as the path with the minimum $\omega$ for battery $i$, as shown in the following equation:
\begin{equation}\label{eq:def_sp}
    SP_i = \mathop{\arg\min}_{p \in P_i} \omega(p),
\end{equation}
where $P_i$ is the set of all paths from $v_1$ to $v_N$ which pass through the directed edge $i$.

The $SP_i$ can be solved by the Dijkstra algorithm.
The Dijkstra algorithm is a graph search method that finds the shortest path between two given nodes in a weighted graph, efficiently solving the single-source shortest path problem.
Assuming that the cathode and anode of battery $i$ are denoted as $v_i^-$ and $v_i^+$ respectively, the path $p$ of battery $i$  can be divided into three segments : $v_1 \rightarrow v_i^-$, $v_i^+ \rightarrow v_N$, and $v_i^- \rightarrow v_i^+$.
The $v_i^- \rightarrow v_i^+$ is the directed edge corresponding to battery $i$. 
With the Dijkstra algorithm, shortest paths for the $v_1 \rightarrow v_i^-$ and $v_i^+ \rightarrow v_N$ can be calculated under the weights given in Equation \ref{eq:weight}, denoted as $SP(v_i^- \rightarrow v_i^+)$ and $SP(v_i^+ \rightarrow v_N)$, respectively.
Finally, the $SP_i$ for battery $i$ is formed by the complete path with $SP(v_1 \rightarrow v_i^-)$, $v_i^- \rightarrow v_i^+$, and $SP(v_i^+ \rightarrow v_N)$.

\subsection{Greedy Algorithm}\label{subsec:greedy_solution}

From the perspective of series/parallel connections, integrating more batteries into the circuit through their shortest paths ($SP$s) results in a larger number of batteries connected in parallel, thereby increasing the total output current of the RBS.
However, conflicts may arise between the $SP$s of different batteries. 
For instance, the $SP$s of two batteries might form a short-circuited RBS structure, which is not allowed. 
To address this issue, a greedy algorithm is employed to incorporate as many $SP$s as possible while satisfying the reconstruction requirements.

The algorithm, as illustrated in Figure \ref{fig:flowchart}, can be summarized as follows, with the corresponding pseudo-code presented in Algorithm \ref{alg:greedy}.
First, the shortest paths ($SP$s) are obtained using Equations \ref{eq:weight} and \ref{eq:def_sp} in conjunction with Dijkstra Search. 
Next, the matrix $\bm{A}$ is calculated using Equation \ref{eq:A}, and the initial $N_{set}$ is set to $N_b$. 
The algorithm iteratively checks different combinations of $c_b$ batteries from $N_b$ and updates $N_{set}$ using a dichotomy method until convergence is reached. 
For each combination, the algorithm constructs an effective solution if possible, and calculates the currents $I_o$ and $\bm{I}_b$ using Equations \ref{eq:I_o} and \ref{eq:I_b}. 
If the maximum current $\bm{I}_b$ is less than or equal to $I_m$, the $\eta$ is calculated using Equation \ref{eq:eta}, and the maximum $\eta$ is updated accordingly. 
Finally, the algorithm outputs the maximum $\eta$ once $N_{set}$ converges.

\tikzset{
  meta box/.style={draw, black, very thick, text centered, },
  punkt/.style={meta box, rectangle, rounded corners, inner sep=.25em, minimum height=2em, minimum width=4em, align=center, text width=10em },
  round/.style={meta box, circle, minimum size=0, inner sep=0pt, outer sep=0pt },
  every fit/.style={draw, thick, dashed, gray, inner xsep=.5em, inner ysep=.75em }
}
\begin{figure}
\begin{tikzpicture}[font=\small, node font=\small, node distance=1.5em]
    \node[punkt]     (input) {Input: RBS structure};
    \node[punkt, below=0.3of input] (get_SP) {get $SP$s by Equations \ref{eq:weight} and \ref{eq:def_sp} and Dijkstra Search};
    \node[punkt, below=0.3of get_SP] (get_A) {get $\bm{A}$ from Equation \ref{eq:A}};
    \node[punkt, below=0.3of get_A] (get_Nset) {init $N_{set}=N_b$ };
    \node[punkt, below=0.3of get_Nset] (get_cb) {get $c_b$s by combinating $N_{set}$ batteries from $N_b$};
    \node[draw, diamond, aspect=2, below=0.3of get_cb] (is_check_all_cb) {are all $c_b$s checked?};
    \node[draw, diamond, aspect=2, right=1of is_check_all_cb] (is_Nset_converged) {is $N_{set}$ converged?};
    \node[punkt, above=of is_Nset_converged] (reset_Nset) {reset $N_{set}$ by dichotomy};
    \node[punkt, text width=15em, below=0.3of is_check_all_cb] (get_Xs) {
        select an unchecked $c_b$, and get its $\bm{X}_m$ by \\ 
        if switch $j$ $\in \bigcup_{i\in c_b}SP_i$:\\
        $\bm{X}[j]=1$ else $0$};
    \node[punkt, below=0.3of get_Xs] (get_Yn) {get $\bm{Y}_n$ by Equation \ref{eq:Yn}};
    \node[draw, diamond, aspect=2, below=0.3of get_Yn] (is_Yn_invertible) {is $\bm{Y}_n$ invertible?};
    \node[punkt, right=1.3of is_Yn_invertible] (construct) {construct an effective solution};
    \node[punkt, below=0.3of is_Yn_invertible] (get_I) {get $I_o$ and $\bm{I}_b$ by Equations \ref{eq:I_o} and \ref{eq:I_b}};
    \node[draw, diamond, aspect=2, below=0.3of get_I] (is_leq_Im) {is $\max \bm{I}_b \leq I_m$?};
    \node[punkt, right=1.3of is_leq_Im] (drop_eta) {drop this $\eta$};
    \node[punkt, below=0.3of is_leq_Im] (get_eta) {get $\eta$ by Equation \ref{eq:eta}};
    \node[punkt, below=0.3of get_eta] (update_max_eta) {update $\max \eta$};
    \node[punkt, right=1of is_Nset_converged] (output) {Output: $\max \eta$};
    \node[round,left=1.5of update_max_eta](point1){};

    \graph{
      (input) -> (get_SP) -> (get_A) -> (get_Nset) -> (get_cb) -> (is_check_all_cb) ->["No"] (get_Xs) -> (get_Yn) -> (is_Yn_invertible) ->["Yes"] (get_I) -> (is_leq_Im) ->["Yes"] (get_eta) -> (update_max_eta);
      (is_check_all_cb) ->["Yes"] (is_Nset_converged) ->["No"] (reset_Nset) -> (get_cb);
      (is_Yn_invertible) ->["No"] (construct) ->[to path={|- (\tikztotarget)}] (get_I);
      (is_leq_Im) ->["No"] (drop_eta) ->[to path={|- (\tikztotarget)}] (update_max_eta);
      (is_Nset_converged) ->["Yes"] (output);
      (update_max_eta) -- (point1) ->[to path={|- (\tikztotarget)}] (is_check_all_cb);
    };
\end{tikzpicture}
\caption{The computational flowchart of the MAC for a given RBS.}\label{fig:flowchart}
\end{figure}

\section{Case Study}

\subsection{Structures}

\begin{figure}[htbp]
    \centering
    \begin{subfigure}[b]{0.2\textwidth}
        \includegraphics[width=\textwidth]{stru-L-origin.png}
        \caption{}
        \label{fig:study-stru-Lawson}
    \end{subfigure}
    \hspace{0.02\textwidth}
    \begin{subfigure}[b]{0.4\textwidth}
        \includegraphics[width=\textwidth]{stru-V-origin.png}
        \caption{}
        \label{fig:study-stru-Visairo}
    \end{subfigure}
    \hspace{0.02\textwidth}
    \begin{subfigure}[b]{0.31\textwidth}
        \includegraphics[width=\textwidth]{stru-my-origin.png}
        \caption{}
        \label{fig:study-stru-my}
    \end{subfigure}
    \caption{The 4-battery RBS structures proposed by (a)Lawson\cite{lawsonSoftwareConfigurableBattery2012}, (b)Visairo\cite{visairoReconfigurableBatteryPack2008} and (c)this paper.}
\end{figure}

Currently, two types of RBS structures have been proposed by Visairo et al. \cite{visairoReconfigurableBatteryPack2008} and Lawson et al. \cite{lawsonSoftwareConfigurableBattery2012}, both of which have been \replaced{practically used}{applied in practice}. 
The primary goal of Visairo's structure (Figure \ref{fig:study-stru-Visairo}) was to achieve dynamic adjustment of RBS output \added{power}; however, the isolation of unhealthy batteries was not sufficiently addressed \added{in their work}. 
\deleted{When batteries need to be isolated in the RBS of Visairo's structure, the methods for isolating them and the subsequent changes in RBS output warrant further investigation.}
Lawson et al. \replaced{designed the RBS structure shown in Figure \ref{fig:study-stru-Lawson} for the purpose of battery isolation}{conducted research on battery isolation in RBS and specifically designed the structure shown in Figure \ref{fig:study-stru-Lawson}}. 
This structure has the advantage of easily isolating batteries, but it cannot dynamically adjust the output current of RBS. 
Based on the structures of Visairo and Lawson, this paper presents a new structure, as shown in Figure \ref{fig:study-stru-my}\deleted{, which combines the advantages of both}.
By integrating the Visairo RBS structure into the Lawson RBS structure, the new structure not only allows the flexibility to switch the batteries between series, parallel, and mixed series-parallel modes, but also easily enables the isolation of highly degraded batteries from the RBS.
And their variations in output current under battery isolation conditions will be studied.
This RBS structure will be used to validate the effectiveness of the proposed method for calculating the MAC, and be compared with the Lawson's and Visairo's structure to illustrate its advantage on battery isolation.

\subsection{Result}

As shown in Figure \ref{fig:study-stru-my}, the new RBS structure consists of 4 batteries and 19 switches. 
The corresponding directed graph is depicted in Figure \ref{fig:study-dirgraph-my}, which is composed of of a total of 18 nodes and 43 edges. 
Batteries $B_1$, $B_2$, $B_3$, and $B_4$ are denoted by green directed edges in the graph, while the 19 switches are represented by gray directed edges with bi-directional arrows. 
The external electrical load is treated as a directed edge from the cathode of the RBS (i.e., node 18) to the anode (i.e., node 1), as indicated by the blue directed edge in the graph.
Utilizing Equation \ref{eq:weight} and the Dijkstra algorithm, the $SP$s of the four batteries in the RBS structure of Figure \ref{fig:study-stru-my} are highlighted by red in Figures \ref{fig:sp1}-\ref{fig:sp4}.
Finally, the MAC calculation results of the structure in Figure \ref{fig:study-stru-my} are shown as Table \ref{tab:study-results-my} and Figure \ref{fig:study-results-my}, obtained by the greedy algorithm \ref{alg:greedy}.
Table \ref{tab:study-results-my} contains the switches states, the output current $I_o$, battery current $\bm{I}_b$ and ratio $\eta$ of the RBS structure with all batteries in good health when the RBS output reaches the MAC.
Figure \ref{fig:study-results-my} presents the corresponding circuit, with the red highlight indicating that current is flowing through the respective branches.

\begin{figure}[htbp]
    \centering
    \begin{subfigure}[b]{0.28\textwidth}
        \includegraphics[width=\textwidth]{ef-topo.png}
        \caption{}
        \label{fig:study-dirgraph-my}
    \end{subfigure}
    \hspace{0.05\textwidth}
    \begin{subfigure}[b]{0.28\textwidth}
        \includegraphics[width=\textwidth]{ef-sp1.png}
        \caption{}
        \label{fig:sp1}
    \end{subfigure}
    \hspace{0.05\textwidth}
    \begin{subfigure}[b]{0.28\textwidth}
        \includegraphics[width=\textwidth]{ef-sp2.png}
        \caption{}
        \label{fig:sp2}
    \end{subfigure}
    \\
    \begin{subfigure}[b]{0.28\textwidth}
        \includegraphics[width=\textwidth]{ef-sp3.png}
        \caption{}
        \label{fig:sp3}
    \end{subfigure}
    \hspace{0.05\textwidth}
    \begin{subfigure}[b]{0.28\textwidth}
        \includegraphics[width=\textwidth]{ef-sp4.png}
        \caption{}
        \label{fig:sp4}
    \end{subfigure}
    \hspace{0.05\textwidth}
    \begin{subfigure}[b]{0.28\textwidth}
        \includegraphics[width=\textwidth]{ef-mac.png}
        \caption{}
        \label{fig:study-results-my}
    \end{subfigure}
    \caption{
        For the RBS structure in Figure \ref{fig:study-stru-my}, 
        (a)its directed graph 
        and the $SP$s (highlighted in red) of battery (b)$B_1$, (c)$B_2$, (d)$B_3$, (e)$B_4$.
        (f)The circuit of the RBS with its output reaching the MAC.
        }
    \label{fig:all-results-my}
\end{figure}

\begin{table}[htbp]
  \centering
    \caption{MAC Calculating result of the 4-battery RBS structure in Figure \ref{fig:study-stru-my}.}
    \begin{tabular}{cc}
    \toprule
        Structure & Figure \ref{fig:study-stru-my} with 4 batteries and 19 switches  \\
    \midrule
    Switch ON & $S_1$,$S_3$,$S_5$,$S_6$,$S_8$,$S_9$,$S_{10}$,$S_{12}$,$S_{18}$,$S_{19}$ \\
    $I_o$ & $2u_b/(2R_o+r_b)$ \\
    $\bm{I}_b$ & $[u_b/(2R_o+r_b),u_b/(2R_o+r_b),0,0]$ \\
    \added{$\max$}$\eta$     & 2 \\
    \bottomrule
    \end{tabular}
  \label{tab:study-results-my}
\end{table}

Similarly, the MAC calculation results of the structures in Figures \ref{fig:study-stru-Lawson} and \ref{fig:study-stru-Visairo} are shown as Table \ref{tab:study-results-Lawson} and Table \ref{tab:study-results-Visairo}, respectively.
\added{
In order to verify and compare the results from the greedy algorithm, we also utilized a brute force algorithm that go through all possible switch states to calculate the MAC of the same three RBS. 
For a given RBS structure with $N_s$ switches, the final $\max$ $\eta$ is the maximum of $\eta$s from all $2^{N_s}$ reconfigured structures.
The final results are the same as the results shown in Tables \ref{tab:study-results-my}-\ref{tab:study-results-Visairo}.
It is worth noting that the method used the greedy algorithm only calculated 7, 11, and 1 reconfigured structures for the RBS structure in Figures \ref{fig:study-stru-my}, \ref{fig:study-stru-Lawson}, \ref{fig:study-stru-Visairo}, respectively. 
While for the same RBS, the method counted all possible switch states computed $2^{19}$, $2^{15}$, and $2^{13}$ structures, respectively.
}

\begin{table}[htbp]
  \centering
    \caption{MAC Calculating result of the 4-battery RBS structure in Figure \ref{fig:study-stru-Lawson}.}
    \begin{tabular}{cc}
    \toprule
        Structure & Figure \ref{fig:study-stru-Lawson} with 4 batteries and 15 switches  \\
    \midrule
    Switch ON & $S_1$,$S_3$,$S_5$,$S_7$,$S_{10}$,$S_{13}$,$S_{14}$,$S_{15}$ \\
    $I_o$ & $u_b/(R_o+r_b)$ \\
    $\bm{I}_b$ & $[u_b/(R_o+r_b),0,0,0]$ \\
    \added{$\max$}$\eta$     & 1 \\
    \bottomrule
    \end{tabular}
  \label{tab:study-results-Lawson}
\end{table}

\begin{table}[htbp]
  \centering
    \caption{MAC Calculating result of the 4-battery RBS structure in Figure \ref{fig:study-stru-Visairo}.}
    \begin{tabular}{cc}
    \toprule
        Structure & Figure \ref{fig:study-stru-Visairo} with 4 batteries and 13 switches  \\
    \midrule
    Switch ON & $S_1$,$S_2$,$S_3$,$S_4$,$S_5$,$S_9$,$S_{10}$,$S_{11}$,$S_{12}$,$S_{13}$ \\
    $I_o$ & $4u_b/(4R_o+r_b)$ \\
    $\bm{I}_b$ & $[u_b/(4R_o+r_b),u_b/(4R_o+r_b),u_b/(4R_o+r_b),u_b/(4R_o+r_b)]$ \\
    \added{$\max$}$\eta$     & 4 \\
    \bottomrule
    \end{tabular}
  \label{tab:study-results-Visairo}
\end{table}

Furthermore, the RBS \replaced{with}{under the scenario of} isolated batteries is taken into consideration and calculated. 
The MAC calculation results for the three structures under study, with varying numbers of isolated batteries, are presented in Table \ref{tab:isolated_mac}. 
Figures \ref{fig:my-isolated-1}-\ref{fig:my-isolated-3} illustrate the corresponding switch control schemes for the new structure proposed in this paper under different isolated battery conditions.
\deleted{The characteristics of these three structures in the context of battery isolation will be discussed in the next subsection.}

\begin{table}[htbp]
    \centering
    \caption{
      The variation of MAC with the number of isolated batteries for different RBS structures, including the structure proposed by Lawson et al., Visairo et al. , and the structure proposed in this paper.
      }
      \label{tab:isolated_mac}
      \begin{tabular}{cccc}
      \toprule
      \multirow{2}[4]{*}{number of isolated batteries} & \multicolumn{3}{c}{$\eta$ of RBS structure} \\
  \cmidrule{2-4}          & our  & Visairo's  & Lawson's  \\
      \midrule
      0     & 2     & 4     & 1 \\
      1     & 2     & 3     & 1 \\
      2     & 2$^{\mathrm{a}}$ or 1$^{\mathrm{b}}$ & 2     & 1 \\
      3     & 1     & 1     & 1 \\
      \bottomrule
      \end{tabular}
      \\
      \footnotesize{$^{\mathrm{a}}$ isolate two batteries within the same substructure, as shown in Figure \ref{fig:my-isolated-2b}}\\
      \footnotesize{$^{\mathrm{b}}$ isolate one battery in each of the two substructures, as shown in Figure \ref{fig:my-isolated-2w}}
  \end{table}

  \begin{figure}[htbp]
      \centering
      \begin{subfigure}[b]{0.31\textwidth}
          \includegraphics[width=\textwidth]{my-isolated-1.png}
          \caption{}
          \label{fig:my-isolated-1}
      \end{subfigure}
      \hspace{0.02\textwidth}
      \begin{subfigure}[b]{0.31\textwidth}
          \includegraphics[width=\textwidth]{my-isolated-2b.png}
          \caption{}
          \label{fig:my-isolated-2b}
      \end{subfigure}
      \\
      \begin{subfigure}[b]{0.31\textwidth}
          \includegraphics[width=\textwidth]{my-isolated-2w.png}
          \caption{}
          \label{fig:my-isolated-2w}
      \end{subfigure}
      \hspace{0.02\textwidth}
      \begin{subfigure}[b]{0.31\textwidth}
          \includegraphics[width=\textwidth]{my-isolated-3.png}
          \caption{}
          \label{fig:my-isolated-3}
      \end{subfigure}
      \caption{
          The circuit states of MACs when isolating (a)one, (b)two(best case), (c)two(worst case) and (d)three batteries for the structure in Figure \ref{fig:study-stru-my}.
          }
  \end{figure}

\subsection{Discussion}

\replaced{As shown in Figure \ref{fig:all-results-my} and Table \ref{tab:study-results-my}}{In this subsection, we firstly discuss the correctness of the results presented in Figure \ref{fig:all-results-my} and Table \ref{tab:study-results-my}}.
When $B_1$ and $B_2$ or $B_3$ and $B_4$ are connected in parallel, the RBS can output the maximum current, which is $\eta=2$, i.e., twice the current output of a single battery in RBS. 
Adding more batteries to the main circuit can only form a series structure and will not improve the MAC. 
Therefore, the switches state given in Table \ref{tab:study-results-my} can make the RBS output current reach the maximum.
\added{The results of MAC obtained by going through all reconfigured structures is identical to the one by the greedy algorithm. 
This further validates the algorithm we proposed.}

\added{
From the literature search we have conducted, no formal report on the algorithm for MAC in RBS has been found yet.
Therefore, the time complexity of the proposed algorithm is analyzed and compared with the algorithm that goes through all reconfigured structures to demonstrate the advantage of this algorithm.
Assuming that a RBS has $N_b$ batteries and $N_s$ switches, and the corresponding directed graph has $N$ nodes, it requires $2^{N_s}$ iterations to traverse all reconfigured structures.
The calculation for each reconfigured structure by Equations \ref{eq:I_o}-\ref{eq:eta} requires matrix inversion and matrix multiplication, with a time complexity of $O(N^3+2N^2N_b+N^2N_s+NN^2_b)$.
Therefore, the time complexity of the brute force algorithm is $O(2^{N_s}(N^3+2N^2N_b+N^2N_s+NN^2_b))$.
The greedy algorithm proposed in this paper requires find the $SP$ for each battery, which requires $N_b$ iterations.
Each $SP$ can be obtained by couple Dijkstra algorithms.
Therefore, the total time complexity of calculating all $SP$s is $O(2N_b(N_b+2N_s)\log N)$.
According to the Appendix \ref{alg:greedy}, the RBS can reconfigure $C^{N_{set}}_{N_b}$ structures by selecting $N_{set}$ batteries from $N_b$ batteries, which is $\sum^{N_b}_{N_{set}=1}C^{N_{set}}_{N_b}/N_b \approx 2^{N_b}/N_b$ on average.
Hence, with bisection method, the time complexity of the greedy algorithm is $O(2^{N_b}/N_b(N^3+2N^2N_b+N^2N_s+NN^2_b)\log N_b+2N_b(N_b+2N_s)\log N)$, i.e., $O(2^{N_b}/N_b(N^3+2N^2N_b+N^2N_s+NN^2_b)\log N_b)$.
Based on currently proposed RBS structures\cite{ciNovelDesignAdaptive2007,alahmadBatterySwitchArray2008,kimDependableEfficientScalable2010b,kimBalancedReconfigurationStorage2011a,taesickimSeriesconnectedSelfreconfigurableMulticell2012a,6843711}, the number of batteries $N_b$, switches $N_s$, and nodes $N$ have the following quantitative relationships: $N_s \approx (3\sim 5)N_b$, $N \approx N_s$. 
After simplifying, the time complexity of the greedy algorithm is $O(2^{N_b}N_s^2\log N_b)$, while it is $O(2^{N_s}N_s^3)$ for brute force algorithm.
Therefore, it is reasonable to believe that as the size of the RBS increases, especially the number of the swithes, the greedy algorithm will have an advantage over the algorithm that goes through all reconfigured structures.
This can be comfirmed from the number of structures required for MAC determination in the previous subsection. 
Compared to the brute force algorithm, the efficiency of the method based on the greedy algorithm has been improved by 3000 to 75000 times, which is theoretically $N_s 2^{N_s - N_b} \log N_b$ times according to the above time complexity analysis.
Among the three RBS structures, the highest is the RBS structure with 19 switches (Figure \ref{fig:study-stru-my}).
This benefit from two key points:
(1) The $SP$s guide the RBS to reconfigure reasonable structures, rather than blindly going through all possible structures. This reduces the factor in complexity from $2^{N_s}$ to $2^{N_b}$, which is the main reason for the improvement in efficiency;
(2) The bisection method further accelerates this process.
However, the greedy algorithm proposed in this paper still contains exponential terms in the time complexity, which means it may not be able to handle extremely large-scale RBS structures.
}

It is important to note that \deleted{when solving for MAC,} $\eta$ is used as the objective function instead of $I_o$ \added{in solving MAC}. 
This choice makes the result of MAC more reasonable. 
As shown in Table \ref{tab:study-results-my}, $I_o$ and $\bm{I}_b$ are functions of $R_o$, $u_b$, and $r_b$. 
\replaced{However, when}{If} $I_o$ \replaced{was}{were} used as the objective function, even for the same RBS structure, the MAC \replaced{solution}{result} and corresponding switches state could change due to different external electrical appliances.
It would increase the difficulty and uncertainty in RBS structure design. 
\replaced{In order to eliminate the influences of such a problem, $\eta$, which is the ratio of $I_o$ and $\max\bm{I}_b$, is adopted as the objective function in our research.}{In contrast, by using $\eta$ as the objective function, which is defined as the ratio of $I_o$ and $\max\bm{I}_b$, the influence of these factors on the results can be eliminated.}
$\eta$ solely reflects the maximum output current capability of the RBS structure. 
Assuming that the maximum allowed current of batteries in the RBS is $I_m$, the maximum output current of the RBS structure can be calculated as $\eta I_m$ by determining the $\eta$ of the structure. 
\deleted{Therefore, compared to $I_o$, $\eta$ is more suitable for structure design.}

The method proposed in this paper is significant for the design of \deleted{next-generation} RBSs in the following aspects.
Most of the currently proposed RBS structures\cite{ciNovelDesignAdaptive2007,alahmadBatterySwitchArray2008,kimDependableEfficientScalable2010b,kimBalancedReconfigurationStorage2011a,taesickimSeriesconnectedSelfreconfigurableMulticell2012a,6843711} exhibit simple topological characteristics, and the calculation of MACs is relatively straightforward, even intuitive.
However, these simple structures do not always fully satisfy the requirements of complex applications, such as dynamically adapting the circuit to variable and random operating conditions, and actively equalizing differences among the batteries in the RBS.
Moreover, \replaced{isolation of}{isolating} the batteries disrupts the original regularity and symmetry of the topology, which complicates the otherwise simple structure, and the maximum output current of the system becomes more challenging to obtain.
\replaced{On the contrary}{Owing to the advantages of pervasiveness and automation}, the proposed method can be employed to calculate the MAC of arbitrary RBS structures, \replaced{especially for these above complex and flexible RBS structures}{which helps to address the aforementioned issues and paves the way for more complex and flexible RBS structure design}.

To illustrate this point, the MACs of \deleted{the} three RBS structures mentioned above are calculated after \replaced{one and more}{the} batteries are isolated, as shown in Table \ref{tab:isolated_mac}. 
Specifically, for the structure presented in Figure \ref{fig:study-stru-my}, the corresponding circuit states of MACs \replaced{of}{when} isolating \replaced{one to three}{different numbers of} batteries are depicted in Figures \ref{fig:my-isolated-1}-\ref{fig:my-isolated-3}. 
This structure has two cases of isolating two batteries: 
one is to isolate two batteries within the same substructure (Figure \ref{fig:my-isolated-2b}), in which case $\eta=2$; the other is to isolate one battery in each of the two substructures (Figure \ref{fig:my-isolated-2w}), in which case $\eta=1$. 
From the results \added{shown in Figure \ref{fig:my-isolated-1}-\ref{fig:my-isolated-3}}, it can be observed that the proposed method provides reasonable outcomes for isolating batteries with any number and position.

Furthermore, the \deleted{performance of} output current for the three RBS \replaced{with isolated}{when isolating} batteries is also shown in Table \ref{tab:isolated_mac}. 
For the structure proposed by Lawson et al., the MAC \replaced{is independent on the number of isolated batteries}{remains the same as that without isolated battery cells, i.e., $\eta=1$, when the number of isolated battery cells increases, until all the cells in the RBS are isolated}.
\added{However,} \replaced{f}{F}or Visairo's structure, the MAC decreases \replaced{with the increasing number of isolated batteries}{as the number of isolated battery cells increases, until $\eta=0$}. 
\replaced{Nevertherless}{In contrast}, the MAC of the structure proposed in this work is positioned between \replaced{these above}{the} two structures. 
This indicates that the structure proposed in this paper, compared to Lawson's structure, has a larger MAC under the same number of batteries, \replaced{and exhibits}{which means} a wider output current regulation range. 
\deleted{On the other hand, by simply changing the states of $S_2$, $S_4$, $S_{11}$, and $S_{12}$ in the conversion structure, this structure can address the majority of battery isolation scenarios, whereas Visairo's structure requires specific battery targeting and switch control.}
\deleted{In summary, the structure proposed in this paper has the advantages of both Lawson's and Visairo's structures.}

\section{Conclusion}

This paper proposes a pervasive and automatical method for computing the MAC of \replaced{a}{the} given RBS.
\added{
The method is implemented by a greedy algorithm combined with a directed graph model which considers the voltage, internal resistance, and maximum allowable current of the batteries, as well as the external load.
The main advantage of this method is its ability to calculate the MAC of RBSs with arbitrary structures.
Even in the scenarios with random isolated batteries, the method remains effective.
Compared with the brute force algorithm, the proposed method has higher computational efficiency under the same calculation results.
This is achieved by two key points when constructing the greedy algorithm: 
(1) Calculate the shortest path of each battery by Dijkstra algorithm to obtain the $SP$ of each battery; 
(2) determine the maximum number of available batteries by bisection method to reduce the computational complexity.
This method helps to fully tap the current output potential of the RBS, guide the RBS structure design and optimization in the design stage, and assist in evaluating the current overload risk of the system in practical applications.
}
\deleted{The method is implemented by a greedy algorithm combined with a directed graph model , whose effectiveness is tested on a novel and complex RBS structure.
The method remains effective for the application scenario of RBS battery isolation and demonstrates that the novel structure has the advantage on flexible output current and convenient battery isolation.
Future research could focus on developing new indictors to evaluate the performance of the RBS with the currents and voltages obtained by the method, as well as modifying the equivalent model of the battery to allow for more accurate simulations of the RBS, including transient analysis.}

\section{Appendix}

\begin{algorithm}
    \caption{Get the max available currents of a certain RBS}\label{alg:greedy}
    \KwData{Directed graph model $G(V,E)$ of the RBS}
    \KwResult{$\max \eta$}
    \For{$i \in E_b$}{
        $P_i \leftarrow \{path| \text{starts at $v_1$ and ends at $v_n$} \}$\;
        $SP_i \leftarrow p_i \text{ which has the minimum}~\omega(p_i)~\text{among all}~p_i \in P_i. $
    }
    get $\bm{A}$ by Equation \ref{eq:A}\;
    \While{not yet determine $\max \eta$ }
    {
        $N_{set} \leftarrow \text{number of setected $SP$s calculated by dichotomy}$\;
        $C_b    \leftarrow \text{set of all combinations of $N_{set} $~batteries from $N_b$}$\;
        \For{$c_b \in C_b$}{
            $\bm{x}_s \leftarrow \text{list of all switches' state: $x_s[j]=1$ if $ j \in \bigcup_{i\in c_b}SP_i $ else 0}$\;
            $\bm{X} \leftarrow diag[1,1,\cdots,1,\bm{x}_s] $\;
            get $\bm{Y}_n$ by Equation \ref{eq:Yn}\;
            \eIf{$\bm{Y}_n$ is invertible}{
              \added{pass}
            }{construct an effective solution}
            get $I_o$ by Equation \ref{eq:I_o}\;
            get $\bm{I}_b$ by Equation \ref{eq:I_b}\;
            \eIf{$\max(\bm{I}_b)\leq I_m$}{
                $\eta \leftarrow I_o/\max(\bm{I}_b)$\;
            }{break}
        }
    }
\end{algorithm}

\section*{Acknowledgments}

\subsection*{Author Contributions}

B. Xu conceived the main idea, formulated the overarching research goals and aims, designed the algorithm, and reviewed and revised the manuscript.
G. Hua developed and analyzed the model, implemented the code and supporting algorithms, and wrote the initial draft.
C. Qian provided critical review, commentary, and revisions.
Q. Xia contributed to shaping the research, analysis, and manuscript.
B. Sun conducted the research and investigation process.
Y. Ren secured the funding and supervised the project.
Z. Wang verified the results and provided necessary resources.

\subsection*{Funding}

This work was supported by the National Natural Science Foundation of China (NSFC, No.52075028).

\subsection*{Conflicts of Interest}

The authors declare that there is no conflict of interest regarding the publication of this article.

\subsection*{Data Availability}

This work does not require any data to be declared or publicly disclosed.
\begin{thebibliography}{10}

\bibitem{alahmadBatterySwitchArray2008}
Mahmoud Alahmad, Herb Hess, Mohammad Mojarradi, William West, and Jay Whitacre.
\newblock Battery switch array system with application for {{JPL}}'s
  rechargeable micro-scale batteries.
\newblock {\em Journal of Power Sources}, 177(2):566--578, March 2008.

\bibitem{chenSneakCircuitTheory2021}
Si-Zhe Chen, Yule Wang, Guidong Zhang, Le~Chang, and Yun Zhang.
\newblock Sneak {{Circuit Theory Based Approach}} to {{Avoiding Short-Circuit
  Paths}} in {{Reconfigurable Battery Systems}}.
\newblock 68(12):12353--12363, 2021.

\bibitem{choCommercialResearchBattery2015}
Jaephil Cho, Sookyung Jeong, and Youngsik Kim.
\newblock Commercial and research battery technologies for electrical energy
  storage applications.
\newblock {\em Progress in Energy and Combustion Science}, 48:84--101, June
  2015.

\bibitem{ci2007novel}
Song Ci, Jiucai Zhang, Hamid Sharif, and Mahmoud Alahmad.
\newblock A novel design of adaptive reconfigurable multicell battery for
  power-aware embedded networked sensing systems.
\newblock In {\em IEEE GLOBECOM 2007-IEEE Global Telecommunications
  Conference}, pages 1043--1047. IEEE, 2007.

\bibitem{ciNovelDesignAdaptive2007}
Song Ci, Jiucai Zhang, Hamid Sharif, and Mahmoud Alahmad.
\newblock A {{Novel Design}} of {{Adaptive Reconfigurable Multicell Battery}}
  for {{Power-Aware Embedded Networked Sensing Systems}}.
\newblock In {\em {{IEEE GLOBECOM}} 2007-2007 {{IEEE Global Telecommunications
  Conference}}}, pages 1043--1047, November 2007.

\bibitem{desiqueiraControlStrategySmooth2021}
Luanna Maria~Silva {de Siqueira} and Wei Peng.
\newblock Control strategy to smooth wind power output using battery energy
  storage system: {{A}} review.
\newblock {\em Journal of Energy Storage}, 35:102252, March 2021.

\bibitem{10.1145/321694.321699}
Jack Edmonds and Richard~M. Karp.
\newblock Theoretical improvements in algorithmic efficiency for network flow
  problems.
\newblock {\em J. ACM}, 19(2):248–264, apr 1972.

\bibitem{9209774}
Jan Engelhardt, Tatiana Gabderakhmanova, Gunnar Rohde, and Mattia Marinelli.
\newblock Reconfigurable stationary battery with adaptive cell switching for
  electric vehicle fast-charging.
\newblock In {\em 2020 55th International Universities Power Engineering
  Conference (UPEC)}, pages 1--6, 2020.

\bibitem{engelhardt2021double}
Jan Engelhardt, Jan~Martin Zepter, Tatiana Gabderakhmanova, Gunnar Rohde, and
  Mattia Marinelli.
\newblock Double-string battery system with reconfigurable cell topology
  operated as a fast charging station for electric vehicles.
\newblock {\em Energies}, 14(9):2414, 2021.

\bibitem{doi:10.1137/0204043}
Shimon Even and R.~Endre Tarjan.
\newblock Network flow and testing graph connectivity.
\newblock {\em SIAM Journal on Computing}, 4(4):507--518, 1975.

\bibitem{fengPropagationMechanismsDiagnosis2019}
Fei Feng, Xiaosong Hu, Lin Hu, Fengling Hu, Yang Li, and Lei Zhang.
\newblock Propagation mechanisms and diagnosis of parameter inconsistency
  within {{Li-Ion}} battery packs.
\newblock {\em Renewable and Sustainable Energy Reviews}, 112:102--113,
  September 2019.

\bibitem{han2021analysis}
Weiji Han, Anton Kersten, Changfu Zou, Torsten Wik, Xiaoliang Huang, and
  Guangzhong Dong.
\newblock Analysis and estimation of the maximum switch current during battery
  system reconfiguration.
\newblock {\em IEEE Transactions on Industrial Electronics}, 69(6):5931--5941,
  2021.

\bibitem{hanNextGenerationBatteryManagement2020a}
Weiji Han, Torsten Wik, Anton Kersten, Guangzhong Dong, and Changfu Zou.
\newblock Next-{{Generation Battery Management Systems}}: {{Dynamic
  Reconfiguration}}.
\newblock {\em IEEE Industrial Electronics Magazine}, 14(4):20--31, December
  2020.

\bibitem{hongwenheStateofChargeEstimationLithiumIon2011}
Hongwen He, Rui Xiong, Xiaowei Zhang, Fengchun Sun, and JinXin Fan.
\newblock State-of-{{Charge Estimation}} of the {{Lithium-Ion Battery Using}}
  an {{Adaptive Extended Kalman Filter Based}} on an {{Improved Thevenin
  Model}}.
\newblock {\em IEEE Transactions on Vehicular Technology}, 60(4):1461--1469,
  May 2011.

\bibitem{heExploringAdaptiveReconfiguration2013}
Liang He, Lipeng Gu, Linghe Kong, Yu~Gu, Cong Liu, and Tian He.
\newblock Exploring {{Adaptive Reconfiguration}} to {{Optimize Energy
  Efficiency}} in {{Large-Scale Battery Systems}}.
\newblock In {\em 2013 {{IEEE}} 34th {{Real-Time Systems Symposium}}}, pages
  118--127, December 2013.

\bibitem{he2014reconfiguration}
Liang He, Linghe Kong, Siyu Lin, Shaodong Ying, Yu~Gu, Tian He, and Cong Liu.
\newblock Reconfiguration-assisted charging in large-scale lithium-ion battery
  systems.
\newblock In {\em 2014 ACM/IEEE International Conference on Cyber-Physical
  Systems (ICCPS)}, pages 60--71. IEEE, 2014.

\bibitem{6843711}
Liang He, Linghe Kong, Siyu Lin, Shaodong Ying, Yu~Gu, Tian He, and Cong Liu.
\newblock Reconfiguration-assisted charging in large-scale {{Lithium-ion}}
  battery systems.
\newblock In {\em 2014 {{ACM}}/{{IEEE International Conference}} on
  {{Cyber-Physical Systems}} ({{ICCPS}})}, pages 60--71, April 2014.

\bibitem{jeevarajanBatterySafetyQualifications2012}
J.~A. Jeevarajan and C.~Winchester.
\newblock Battery {{Safety Qualifications}} for {{Human Ratings}}.
\newblock {\em Interface magazine}, 21(2):51--55, January 2012.

\bibitem{kim2009dynamic}
Hahnsang Kim and Kang~G Shin.
\newblock On dynamic reconfiguration of a large-scale battery system.
\newblock In {\em 2009 15th IEEE Real-Time and Embedded Technology and
  Applications Symposium}, pages 87--96. IEEE, 2009.

\bibitem{kimDependableEfficientScalable2010b}
Hahnsang Kim and Kang~G. Shin.
\newblock Dependable, efficient, scalable architecture for management of
  large-scale batteries.
\newblock In {\em Proceedings of the 1st {{ACM}}/{{IEEE International
  Conference}} on {{Cyber-Physical Systems}}}, {{ICCPS}} '10, pages 178--187,
  {New York, NY, USA}, April 2010. {Association for Computing Machinery}.

\bibitem{taesickimSeriesconnectedSelfreconfigurableMulticell2012a}
Taesic Kim, Wei Qiao, and Liyan Qu.
\newblock A series-connected self-reconfigurable multicell battery capable of
  safe and effective charging/discharging and balancing operations.
\newblock In {\em 2012 {{Twenty-Seventh Annual IEEE Applied Power Electronics
  Conference}} and {{Exposition}} ({{APEC}})}, pages 2259--2264, February 2012.

\bibitem{kimBalancedReconfigurationStorage2011a}
Younghyun Kim, Sangyoung Park, Yanzhi Wang, Qing Xie, Naehyuck Chang, Massimo
  Poncino, and Massoud Pedram.
\newblock Balanced reconfiguration of storage banks in a hybrid electrical
  energy storage system.
\newblock In {\em 2011 {{IEEE}}/{{ACM International Conference}} on
  {{Computer-Aided Design}} ({{ICCAD}})}, pages 624--631, November 2011.

\bibitem{lawsonSoftwareConfigurableBattery2012}
Barrie Lawson.
\newblock A {{Software Configurable Battery}}.
\newblock {\em EVS26 International Battery, Hybrid and Fuel Cell Electric
  Vehicle Symposium}, 2012.

\bibitem{lotfi2017li}
Nima Lotfi, Jie Li, Robert~G Landers, and Jonghyun Park.
\newblock Li-ion battery state of health estimation based on an improved single
  particle model.
\newblock In {\em 2017 American Control Conference (ACC)}, pages 86--91. IEEE,
  2017.

\bibitem{mousavig.VariousBatteryModels2014}
S.M. Mousavi~G. and M.~Nikdel.
\newblock Various battery models for various simulation studies and
  applications.
\newblock {\em Renewable and Sustainable Energy Reviews}, 32:477--485, April
  2014.

\bibitem{pomboHybridPowerSystem2021}
Daniel~V{\'a}zquez Pombo.
\newblock A {{Hybrid Power System}} for a {{Permanent Colony}} on {{Mars}}.
\newblock {\em Space: Science \& Technology}, 2021, January 2021.

\bibitem{schwanbeckInternationalSpaceStation2019}
Eugene Schwanbeck and Penni Dalton.
\newblock International {{Space Station Lithium-ion Batteries}} for {{Primary
  Electric Power System}}.
\newblock In {\em 2019 {{European Space Power Conference}} ({{ESPC}})}, pages
  1--1. {IEEE}, September 2019.

\bibitem{visairoReconfigurableBatteryPack2008}
H.~Visairo and P.~Kumar.
\newblock A reconfigurable battery pack for improving power conversion
  efficiency in portable devices.
\newblock In {\em 2008 7th {{International Caribbean Conference}} on
  {{Devices}}, {{Circuits}} and {{Systems}}}, pages 1--6. {IEEE}, April 2008.

\bibitem{yangUnbalancedDischargingAging2016}
Naixing Yang, Xiongwen Zhang, BinBin Shang, and Guojun Li.
\newblock Unbalanced discharging and aging due to temperature differences among
  the cells in a lithium-ion battery pack with parallel combination.
\newblock {\em Journal of Power Sources}, 306:733--741, February 2016.

\bibitem{yangBatteryEnergyStorage2018}
Yuqing Yang, Stephen Bremner, Chris Menictas, and Merlinde Kay.
\newblock Battery energy storage system size determination in renewable energy
  systems: {{A}} review.
\newblock {\em Renewable and Sustainable Energy Reviews}, 91:109--125, August
  2018.

\bibitem{zhangDevelopmentProspectChinese2021}
Lihua Zhang.
\newblock Development and {{Prospect}} of {{Chinese Lunar Relay Communication
  Satellite}}.
\newblock {\em Space: Science \& Technology}, 2021, January 2021.

\end{thebibliography}

% \bibliographystyle{plain}
% \bibliography{sst_main}

\end{document}