### 2023-09-15 paper/SST/review_response.tex
更正了对 my structure 计算过程中重构结构数的结果，由7改为11

\documentclass[12pt,american]{scrartcl}
\usepackage{babel}
\usepackage[babel]{microtype}
\usepackage[babel]{csquotes}

\usepackage[journal={Space: Science \& Technology},
      manuscript={SPACE-D-23-00082},
      editor={Dr. Tian}]{reviewresponse}

\usepackage[T1]{fontenc}
\usepackage{lmodern}
%\usepackage{newcent}
%\usepackage[scaled]{beramono}

\usepackage[]{biblatex}
\bibliography{2-SST_review.bib}

\usepackage{hyperref}

\title{Maximum Allowable Current Determination of RBS By Using a Directed Graph Model and Greedy Algorithm}
\author{Dr. Cheng Qian, on behalf of all authors}

\begin{document}
\maketitle

% Cover Letter
\include{2-SST_cover_letter}

% Reviewer 1
\reviewer
% Comment 1
\begin{revcomment}
  The authors should explain the most important achievements of the proposed method quantitatively, in the abstract.
\end{revcomment}
\begin{revresponse}

Thanks for the reviewer's valuable feedback. We have carefully considered this suggestion and made the following modifications to address the reviewer's concern:

By introducing the shortest path($SP$) of the battery, the greedy algorithm transforms the enumeration of switch states in the brute force algorithm into the combination of the $SP$s, which greatly increases the efficiency of determining the maximum allowable current (MAC) of reconfigurable battery systems (RBSs). We have also provided a theoretical estimation of the improved efficiency, which is proportional to $N_s 2^{N_s - N_b} \log N_b$, where $N_s$ is the number of switches and $N_b$ is the number of batteries.

Here is the specific modification we made in the abstract:
\begin{changes}
  changes in abstract
\end{changes}

We believe that these modifications provide a more quantitative explanation of the achievements of our proposed method in the abstract.

\end{revresponse}

% Comment 2
\begin{revcomment}
  The literature review in the introduction section is very short, and the related works, especially the works published in recent years, have not been well reviewed and compared, and the conclusions about the existing research gaps have not been presented.
\end{revcomment}
\begin{revresponse}

Dear Reviewer,

We have expanded the literature review in the introduction section to provide a comprehensive overview of the existing RBS structures and related works on structure analysis. 
Although many RBS structures have been proposed for different purposes, such as dynamically adjusting the output voltage, increasing energy utilization efficiency, and improving the system's ability to recover from battery failures, they also bring challenges in design and control of the systems.
Therefore, several works on structure analysis, like the maximum switch current and the short-circuit problem, have been proposed to tackle these challenges recently.
However, determining the MAC of RBSs remains blank accroding to our literature review.
A straightforward method is to enumerate all possible switch states, but the complexity of this method increases exponentially with the number of switches, and has too ineffection to apply.

Here is the specific modification we made in the introduction:
\begin{changes}
  changes in introduction
\end{changes}

Thanks once again for the reviewer's comments, which have greatly improved the quality and comprehensiveness of our manuscript. We believe that the revised introduction section now provides a thorough review of the literature and addresses the existing research gaps in the field of RBS structures.

\end{revresponse}

% Comment 3
\begin{revcomment}
  It is necessary for the authors to clearly state research contribution and achievements as bullet points at the end of the Introduction section.
\end{revcomment}
\begin{revresponse}

We appreciate and accept your suggestion and have added a clear statement of the contributions in the second-to-last paragraph of the Introduction, as shown below:
\begin{changes}
  changes in introduction
\end{changes}

We believe that these additions enhance the clarity of our manuscript.

\end{revresponse}

% Comment 4
\begin{revcomment}
  The authors need to present the complexity of their proposed method and compare it with some other state-of-the-art or successful classic methods.
\end{revcomment}
\begin{revresponse}

Thanks for the reviewer's valuable feedback. We have carefully considered the reviewer's suggestion regarding the complexity of our proposed method and have made the following explanation to address this concern:

We have derived the average time complexity of our proposed greedy algorithm-based MAC determination method to be approximately $O(2^{N_b}N_s^2\log N_b)$, where $N_b$ and $N_s$ are the number of batteries and switches, respectively.

As mentioned in our response to Comment 2, there was a blank in the literature regarding MAC determination methods.
Therefore, a brute force method have been used as a benchmark for comparison, whose time complexity is $O(2^{N_s}N_s^3)$.

Since the number of switches in RBS is typically 3 to 5 times the batteries\cite{ciNovelDesignAdaptive2007,alahmadBatterySwitchArray2008,kimDependableEfficientScalable2010b,kimBalancedReconfigurationStorage2011a,taesickimSeriesconnectedSelfreconfigurableMulticell2012a,6843711}, the method we proposed is theoretically more efficient than the brute force method.
It has been validated by the case study in the manuscript.

The detailed derivation and discussion of the above points have been added to the revised manuscript under the "Discussion".
Here is the specific modification:
\begin{changes}
  changes in discussion
\end{changes}

\end{revresponse}

% Comment 5
\begin{revcomment}
  The authors don't discuss the limitations of the study correctly.
\end{revcomment}
\begin{revresponse}

We are sorry for our negligence of the limitations of the study. As far as we are concerned, although the method we proposed is strongly efficient than the brute force method, it still has a exponential relationship with the number of batteries, which means unsufferable time cost will be caused for systems with large number of batteries.

Here is the specific content we added in the discussion:
\begin{changes}
  changes in discussion
\end{changes}

\end{revresponse}

% Comment 6
\begin{revcomment}
  Some typos should be double check.
\end{revcomment}
\begin{revresponse}

We are very sorry for our incorrect writing. We have carefully checked the manuscript and corrected the typos. Thanks for the valuable feedback.

\end{revresponse}

% Comment 7
\begin{revcomment}
  The author should explain more why solution quality of their proposed approach is much better than the others?
\end{revcomment}
\begin{revresponse}

We understand the reviewer's concern, and we agree that the explanation on the reason why the solution quality of our proposed approach is much better than the others is important to the academic value of the paper. 
However, as metioned in the response to Comment 2,
we have not found any other works on MAC determination of RBSs, which means there is no other solution to compare with.

\end{revresponse}

% Comment 8
\begin{revcomment}
  Authors should mention some novel works in the field in the introduction, specially refer to this 2023 reference: An efficient lightweight algorithm for scheduling tasks onto dynamically reconfigurable hardware using graph-oriented simulated annealing, which uses graph-based method. Mention and refer to it in the introduction section.
\end{revcomment}
\begin{revresponse}

Thanks to the reviewer's suggestion, we have introduced two novel works in the RBS structure analysis field in the introduction, which respectively study the maximum switch current \cite{han2021analysis} and the short-circuit problem \cite{chenSneakCircuitTheory2021} in RBS structures.
We believe that these works represent the innovation and recent advances in the field of RBS structure analysis.

The reviewer also specially mentioned the work published in 2023: An efficient lightweight algorithm for scheduling tasks onto dynamically reconfigurable hardware using graph-oriented simulated annealing.
We have carefully read this paper, and after fully discussion, we finally agreed that this paper is not relevant to our research.
Although this paper also uses a graph-based method, it mainly studies the task scheduling problem in time series, while our research is about the maximum allowable current of RBS structure, which belongs to the field of structure analysis.
Therefore, we believe that it is not appropriate to include this paper in our introduction and will not cite it.

Thanks again for the valuable feedback from the reviewer.

\end{revresponse}

% Comment 9
\begin{revcomment}
  Authors need to explain about the accuracy, sufficiency and reliability of their results? How do they verify and validate the results?
\end{revcomment}
\begin{revresponse}

\end{revresponse}

\end{document}