### 2023-09-15 paper/SST/2-SST_cover_letter.tex
review cover letter (done)

\noindent Dear \editorname,

On behalf of my co-authors, we thank you very much for giving us an opportunity to revise our manuscript, we appreciate editor and reviewer very much for their positive and constructive comments and suggestions on our manuscript entitled \enquote{\thetitle}(ID:\manuscript).

We have studied reviewer's comments carefully and have made revision which marked in red in the paper. We have tried our best to revise our manuscript according to the comments. Specially, this manuscript has been retouched by the professional organization, to further improve the readability. Attached please find the revised version, which we would like to submit for your kind consideration.

We would like to express our great appreciation to you and reviewers for comments on our paper. Looking forward to hearing from you.

Thank you and best regards.

\vspace{1.2em}

\noindent Sincerely,

\vspace{1.5em}

\noindent \theauthor