### 2023-11-10 paper/SST/3-cover_letter.tex
fix and modify cover letter and response

\noindent Dear \editorname \, and the anonymous reviewers,

On behalf of my co-authors, we express our sincere gratitude for providing us with the opportunity to revise our manuscript. We highly value the reviewers' insightful and constructive comments and suggestions on our manuscript titled "\thetitle" (ID:\manuscript).

We have thoroughly examined the reviewers' comments and have made revisions, which are indicated in red and blue within the paper. Please find attached the revised version, which we kindly request you to consider.

We would like to extend our great appreciation to you and the reviewers for their valuable feedback on our paper. We eagerly await your response.

Thank you and best regards.
\begin{flushright}
\theauthor

Beihang University

2023.11.10
\end{flushright}